---
title: "PSM"
date: "2024-02-18"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Load Data
```{r,include=FALSE}
library(tidyverse)
library(MatchIt)
library(lmtest)
library(sandwich)
df = read_csv("https://raw.githubusercontent.com/gckc123/ExampleData/main/smoking_psyc_distress.csv")
```

* sex (0: Female; 1: Male)
* indigenous - Ingigenous status (0: Non-indigenous; 1: indigenous)
* high_school - Education level (0: not finished high school; 1: finished high school)
* partnered - Marital status (0: not partnered; 1: partnered)
* remoteness - region of residence (0: major cities; 1: inner regional; 2: outer regional)
* language - Language background (0: non-English speaking; 1: English speaking)
* smoker - Smoking status (0: non-smoker; 1: smoker)
* risky_alcohol - Risky alcohol use (0: not risky; 1: risky)
* psyc_distress - Psychological distress. Measure ranges from 10 to 50.
* age - Age of the participants

## PSM

```{r}
#Since remoteness is a categorical variable with more than two categories. It is necessary to convert 
#it into a factor variable.
#For other categorical variable with only 2 levels, this is optional if the variable is coded as 0 and 1.
df$remoteness <- factor(df$remoteness, exclude = c("", NA))

#Using the mathcit function from MatchIt to match each smoker with a non-smoker (1 to 1 matching) based on
#sex, indigeneity status, high school completion, marital status (partnered or not),
#region of residence (major cities, inner regional, outer regional), language background (English speaking Yes/No) 
#and risky alcohol drinking (Yes/No)
match_obj <- matchit(smoker ~ sex + indigeneity + high_school + partnered + remoteness + language + risky_alcohol + age,
  data = df, method = "nearest", distance ="glm",
  ratio = 1,
  replace = FALSE)
summary(match_obj)

#plotting the balance between smokers and non-smokers
plot(match_obj, type = "jitter", interactive = FALSE)
plot(summary(match_obj), abs = FALSE)
```

## Output the result

```{r}
#Extract the matched data and save the data into the variable matched_data
matched_data <- match.data(match_obj)

#Run regression model with psychological distress as the outcome, and smoker as the only predictor
#We need to specify the weights - Matched participants have a weight of 1, unmatched participants 
res <- lm(psyc_distress ~ smoker, data = matched_data, weights = weights)
summary(res)
#Test the coefficient using cluster robust standard error
coeftest(res, vcov. = vcovCL, cluster = ~subclass)
#Calculate the confidence intervals based on cluster robust standard error
coefci(res, vcov. = vcovCL, cluster = ~subclass, level = 0.95)

```